Overview and Summary of the First Workshop on Sustainable Software for Science: Practice and Experiences (WSSSPE)

### 2014-02-26 WSSSPE1_report/WSSSPE1_summary_report.tex
fixed a few typos

Bourne's keynote presentation was entitled ``A Recipe for Sustainable
Software,'' developed based on his experiences.  He emphasized that
sustainability for software ``does not just mean more money from
Government'' (see also Section~\ref{sec:cross-cutting}).  Other
factors to consider, he mentioned, include costs of production, ease
of maintainence, community involvement, and distribution channels.

@@

He also had problems with selling research software: the technology
transfer office in his university wanted huge intellectural property
reach through, whereby they would get a share of profits from drugs
developed by pharmaceutical companies who use the software.  He was
aware this was unrealistic but the technology transfer office insisted
for a while. He wants to push a one-click approach for customers to
purchase university-written software.

@@

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as
\cite{NSF_software_vision}. He urged computational scientists to
collabroate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

One paper by Maheshwari et al.~\cite{Maheshwari_WSSSPE} that was
discussed was about the role of technological catalysts in overall e-research
process.  The paper focuses on ``technology catalysts" and their role in modern
scientific research process. Technology catalyst could be defined as a role
played by an individual with a knowhow of technological advancements tasked
with user engagement with a goal of enacting scientific or engineering
applications using suitable tools and techniques to take advantage of
technological capabilities and benefitting applications.

One of the tasks of catalysts is to seek community collaborations for new
applications and user engagement thus benefitting both: science, by effective
running of scientific codes over computational infrastructure and technology,
by conducting research and seeking findings to improve technology.

@@

The Christopherson paper reports on the application of Steps 1-3 to a computational modeling framework developed in the 1990s.
Step 1 was realized as a 2-day, in person specifications meeting and code walk-through. Step 2 was a 5-day hackathon to develop working code.
Step 3 was a 3-day hackathon to refine the code based on new requirements. The team worked on small, low-risk unit of
code. It was challenging, revealed unanticipated obstacles, programmers had to work together, and experimentation was encouraged.

@@

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is especially relevant in the modern era of social computing.
In the context of this paper, a utility software project could be defined as
general purpose software applicable to one or many generic tasks and/or
enabling other software to run by providing a suitable environment. Examples
are collaborative development framework such as \emph{GitHub} and
\emph{Bitbucket}, workflow and scripting frameworks such as \emph{Galaxy},
\emph{Swift}, \emph{Globus} and \emph{VisTrails}, and visualization frameworks
such as \emph{visIT}. Often projects start with specific, small set of
requirements and users in mind but expand to provide more generic functionality
and appeal to a wider user community. Owing to its generic nature, utility
software finds a wider users communities which in turn leads to a higher
participation in development. Consequently, the process becomes user-driven and
self sustaining. Software developed in such scenario often benefits from
community driven development, deployment and maintenance best practices. For
instance the \emph{Galaxy} project has significantly benefitted from user
community participation in its development. It follows agile software
development practices and implements standard practices such as test-driven
development and socialized bug managing practices via \emph{trello}. Galaxy
\emph{histories} and \emph{toolshed} offers easy community sharing of data and
tools further promoting a collaborative environment. The project very closely
follows the guidelines described in~\cite{Carver_WSSSPE} and many
from~\cite{Prlic_WSSSPE}.

>>>>

Bourne's keynote presentation was entitled ``A Recipe for Sustainable
Software,'' developed based on his experiences.  He emphasized that
sustainability for software ``does not just mean more money from
Government'' (see also Section~\ref{sec:cross-cutting}).  Other
factors to consider, he mentioned, include costs of production, ease
of maintenance, community involvement, and distribution channels.

@@

He also had problems with selling research software: the technology
transfer office in his university wanted huge intellectual property
reach through, whereby they would get a share of profits from drugs
developed by pharmaceutical companies who use the software.  He was
aware this was unrealistic but the technology transfer office insisted
for a while. He wants to push a one-click approach for customers to
purchase university-written software.

@@

In conclusion Smith reiterated the importance of establishing
incentive models for open contributions and tool builders, for
example, meaningful metrics and research grants such as
\cite{NSF_software_vision}. He urged computational scientists to
collaborate and share often their research reports, teaching
materials, code, as well as data by attaching proper licenses.

@@

One paper by Maheshwari et al.~\cite{Maheshwari_WSSSPE} that was
discussed was about the role of technological catalysts in overall e-research
process.  The paper focuses on ``technology catalysts" and their role in modern
scientific research process. Technology catalyst could be defined as a role
played by an individual with a know-how of technological advancements tasked
with user engagement with a goal of enacting scientific or engineering
applications using suitable tools and techniques to take advantage of
technological capabilities and benefiting applications.

One of the tasks of catalysts is to seek community collaborations for new
applications and user engagement thus benefiting both: science, by effective
running of scientific codes over computational infrastructure and technology,
by conducting research and seeking findings to improve technology.

@@

The Christopherson paper reports on the application of Steps 1-3 to a computational modeling framework developed in the 1990s.
Step 1 was realized as a 2-day, in person specifications meeting and code walkthrough. Step 2 was a 5-day hackathon to develop working code.
Step 3 was a 3-day hackathon to refine the code based on new requirements. The team worked on small, low-risk unit of
code. It was challenging, revealed unanticipated obstacles, programmers had to work together, and experimentation was encouraged.

@@

\subsection{Utility Software}
Utility software potentially has the broadest community impact of the three
categories. This is especially relevant in the modern era of social computing.
In the context of this paper, a utility software project could be defined as
general purpose software applicable to one or many generic tasks and/or
enabling other software to run by providing a suitable environment. Examples
are collaborative development framework such as \emph{GitHub} and
\emph{Bitbucket}, workflow and scripting frameworks such as \emph{Galaxy},
\emph{Swift}, \emph{Globus} and \emph{VisTrails}, and visualization frameworks
such as \emph{visIT}. Often projects start with specific, small set of
requirements and users in mind but expand to provide more generic functionality
and appeal to a wider user community. Owing to its generic nature, utility
software finds a wider users communities which in turn leads to a higher
participation in development. Consequently, the process becomes user-driven and
self sustaining. Software developed in such scenario often benefits from
community driven development, deployment and maintenance best practices. For
instance the \emph{Galaxy} project has significantly benefited from user
community participation in its development. It follows agile software
development practices and implements standard practices such as test-driven
development and socialized bug managing practices via \emph{trello}. Galaxy
\emph{histories} and \emph{toolshed} offers easy community sharing of data and
tools further promoting a collaborative environment. The project very closely
follows the guidelines described in~\cite{Carver_WSSSPE} and many
from~\cite{Prlic_WSSSPE}.
